An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-03-01 uai2017/paper.tex
Whew, MASSIVE (well ok semi-massive) change for UAI preparation.
We now have a far expanded appendix, revised logistic regression stuff, etc. I still have one LAST LR thing running ... hopefully by Friday we won't have anything else to do except for SMF

@@

% Daniel: use this algorithm style instead
\begin{algorithm}[tb]
\caption{Our acceptance test for MCMC.}
\label{alg:our_algorithm}
\begin{algorithmic}
    \STATE {\bfseries Input:} number of samples $T$, minibatch size $m$, error
        bound $\delta$, pre-computed correction $C_1(X)$ distribution, initial
        sample $\theta_1$.
    \STATE {\bfseries Output:} a chain of $T$ samples
        $\{\theta_1,\ldots,\theta_T\}$.
    \FOR{$t=1$ {\bfseries to} $T$}
        \STATE -Propose a candidate $\theta'$ from proposal
            $q(\theta' | \theta_t)$.
        \STATE -Draw a minibatch of $m$ points $\{x_1^*,\ldots,x_m^*\}$.
        \STATE -Compute $\Delta^*(\theta_t,\theta')$ and sample variance
            $s^2_{\Delta^*}$.
        \STATE -Estimate moments $\mE[|\Lambda_i-\Lambda|]$ and
            $\mE[|\Lambda_i-\Lambda|^3]$ from the sample, and error $\epsilon$
            from Corollary~\ref{cor:our_bound_delta_prime}.
        \WHILE{$s^2_{\Delta^*} \ge 1$ {\bf or} $\epsilon > \delta$}
            \STATE -Draw $m$ more samples to augment the minibatch, update
                $\Delta^*$, $s^2_{\Delta^*}$ and $\epsilon$ estimates.
        \ENDWHILE
        \STATE -Draw $X_{\rm nc} \sim \mathcal{N}(0,1-s^2_{\Delta^*})$ and
            $X_{\rm corr} \sim C_1(X)$.
    \IF{$\Delta^* + X_{\rm nc} + X_{\rm corr} > 0$}
        \STATE -Accept the candidate, $\theta_{t+1} = \theta'$.
    \ELSE
        \STATE -Reject and re-use the old sample, $\theta_{t+1} = \theta_t$.
    \ENDIF
    \ENDFOR
\end{algorithmic}
\end{algorithm}

With this approach, there is no guarantee that $u^* \geq 0$. However, we have
some flexibility in the choice of $\sigma$ in Equation~(\ref{eq:overall_corr_problem}).
As we decrease the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$
grows by the same amount and is in fact the result of convolution with a
Gaussian whose variance is the difference.  Thus as $\sigma$ decreases,
$C_\sigma(X)$ grows and approaches the derivative of a logistic function at
$\sigma = 0$. It retains some weak negative values for $\sigma > 0$ but removal
of those leads to small error. \textbf{Daniel: I moved this to
Appendix~\ref{app:correction} so that we can have more detailed results. For the
main text, let's state the value $\sigma$ and its error. Originally, we said:
Table~\ref{tab:xcorr} shows that the errors between $X_{\rm norm}+X_{\rm corr}$
and $X_{\rm log}$ approach single floating precision (about $10^{-7}$).}
Algorithm~\ref{alg:our_algorithm} describes our procedure. A few points:

@@

@@

@@

We ran MCMC with with our test against the test
from~\cite{icml2014c1_bardenet14} and a simplified version
of~\cite{cutting_mh_2014}. Both methods use repeated testing which must be
discounted to bound overall test accuracy. ~\cite{icml2014c1_bardenet14}
conservatively partitions the test error, while~\cite{cutting_mh_2014} used a
complex dynamic programming heuristic to estimate the overall test error. We did
not implement this heuristic and instead set the individual test error equal to
the overall test error. Thus our results for~\cite{cutting_mh_2014} should be
treated as lower bounds. All methods were initialized with minibatch size 100.
For~\cite{cutting_mh_2014} we increment sizes by 100 and set the tolerance
$\epsilon=0.005$ to control overall test error.
For~\cite{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma =
1.5$ and use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
samples using a random walk proposer with covariance ${\rm diag}(0.15, 0.15)$,
which means the M-H test is responsible for shaping the sample distribution.

Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
samples for the three methods, with darker regions indicating a greater density
of points. There are no obvious differences, so we measure the similarity
between each set of samples and the actual posterior. \textbf{Daniel: I think
this was taking up too much space so I put it in Appendix~\ref{app:gaussians}.
Let's state the result (in Table~\ref{tab:poissons}) and prioritize other stuff
here.  TODO}

Figure~\ref{fig:gauss_mix_2} shows that the new method dominates in
terms of speed and efficiency. The histograms of the (final) minibatch sizes
used each iteration show that our method consumes significantly less data; the
distribution is short-tailed and the mean is 210, more than an order of
magnitude better compared to the other two methods (averages are 15562 and
16857). The sizes correspond to the running times of the methods, since
they all consume linear time in data consumed.

@@

% Daniel: moving this here so figures don't overflow to next pages. We can tune
% the positioning if needed.
\begin{figure*}[t]
	\centering
	\includegraphics[width=1.0\linewidth]{logistic_regression_error_regions.png}
	\caption{
    TODO
    }
	\label{fig:logistic_performance}
\end{figure*}

\begin{figure*}[t]
	\centering
    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist.png}
	\caption{
    Minibatch sizes used for one representative trial in
    Section~\ref{ssec:logistic} using MNIST-13k (analogous to
    Figure~\ref{fig:gauss_mix_2}).
    }
	\label{fig:logistic_histograms_mnist}
\end{figure*}

\begin{figure*}[t]
	\centering
    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist8m.png}
	\caption{
    Minibatch sizes used for one representative trial in
    Section~\ref{ssec:logistic} using MNIST8M-100k (similar to
    Figure~\ref{fig:logistic_histograms_mnist}).
    }
	\label{fig:logistic_histograms_mnist8m}
\end{figure*}

\begin{table}[t]
\caption{Logistic Regression Average Minibatch Size}
\label{tab:logistic}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Method/Data} & \textbf{MNIST-13k} & \textbf{MNIST-100k}  \\
\hline \\
Our Method & $129.8\pm 3.4$ & $2059.9\pm 135.5$ \\
AustereMH  & $1924.3\pm 52.4$ & \textbf{TODO} \\
MHSubLhd   & $10783.4\pm 78.9$ & \textbf{TODO} \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

\textbf{Daniel: TODO this section is badly out of date but I will re-write it
and it should be much better than what we had earlier. Also, define a trial
since I say it in the table and figures. Finally, define AustereMH and MHSubLhd,
and come up with a name for our method.}

We next test logistic regression for the binary classification of 1s versus 7s
on a subset of the MNIST8M dataset, a larger version of
MNIST~\cite{lecun-mnisthandwrittendigit-2010}. We randomly subsampled 450k
training and 192k testing points.  We impose a uniform prior on $\theta$ and
again use a random walk proposer, this time with covariance matrix $0.05I$. The
temperature is set at $T=1000$. We run the three tests for 3000 samples and
again use 100 as the starting minibatch size. Our undiscounted implementation
of~\cite{cutting_mh_2014} used per-test error of $\epsilon=0.05$.
For~\cite{icml2014c1_bardenet14}, we use their recommended discounting factor
value of $\gamma=2.0$. We tried to use the symbolic bound for
$C_{\theta,\theta'}$ for Logistic regression from~\cite{icml2014c1_bardenet14},
but found it to be too high. Instead we use the empirical $C_{\theta,\theta'}$
from the entire dataset and do not add its computation time when considering
runtime, so again our results represent a lower bound for this method.

Figure~\ref{fig:logistic_performance} shows the test log likelihood and
prediction accuracy as a function of the cumulative training points
processed.\footnote{The curves do not span the same length over the x-axis since
the methods consume different amounts of data.} To generate the curves, for each
of the sampled vectors $\theta_t$, $t\in\{1,\ldots,3000\}$, we use $\theta_t$ as
the parameter for logistic regression.  Our minibatch MH test is more efficient
in terms of accuracy, achieving convergence roughly twice as fast compared
to~\cite{cutting_mh_2014} and about 1.5 orders of magnitude faster
than~\cite{icml2014c1_bardenet14}, though its log likelihood results lag
slightly behind the former algorithm during the very early stages of
exploration.

Figure~\ref{fig:logistic_minibatch} shows histograms of minibatch sizes for the
three methods on a log-log scale. With an initial size of 100, we achieve an
average minibatch size of 393, more than two orders of magnitude smaller
than~\cite{icml2014c1_bardenet14}, and almost an order of magnitude faster than
the undiscounted implementation of~\cite{cutting_mh_2014}.

@@

@@

\textbf{Daniel: double check this in case we add a third experiment.}

We have derived an M-H test for minibatch MCMC which approximates full data
tests. We present theoretical results and experimentally show the benefits of
our test on Gaussian mixtures and logistic regression.  Directions for future
work include integrating our algorithm with~\cite{cutting_mh_2014} by applying
both tests each iteration or utilizing the variance reduction techniques
suggested in~\cite{DBLP:conf/icml/ChenG16}.  More elaborate extensions involve
running more experiments on other models such as neural networks and providing a
recipe for how to use our algorithm (following the framework
of~\cite{sgmcmc_2015}).

@@

This appendix is divided into two main parts. Appendix~\ref{app:proofs} provides
the proofs that we omitted from the main text due to space constraints.
Appendix~\ref{app:experiments} provides further details on the correction
distribution derivation and on our three main experiments to assist
understanding and reproducibility.

@@

We first observe that
\[
    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
\]
and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
\begin{equation}
-\epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
\end{equation}
% Daniel: this is the multi-line version, but we should use the full page.
%\begin{align*}
%-\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
%&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
%&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
%\end{align*}
as desired.

@@

%% % Daniel: moved to appendix so we can expand it. Originally we had only two
%% % lines, and it looked pretty awkward. So let's add more. Note: this also
%% % depends on whether we use [-10,10] for the X_c range, or [-20,20], etc...
\begin{table}[t]
\caption{Errors ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm
log}$, with $N=4000$ (top row) and $N=2000$ (bottom row).}
\label{tab:xcorr}
\vskip 0.15in
\begin{center}
\begin{small}
%% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=2000$ & $\sigma=0.8$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 2.6e-3  \\
10     & 4.0e-4  \\
1      & 6.7e-5  \\
0.1    & 1.4e-5  \\
0.01   & \textbf{5.0e-6}  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=2000$ & $\sigma=0.9$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 3.3e-3  \\
10     & 6.4e-4  \\
1      & 1.6e-4  \\
0.1    & \textbf{1.3e-4}  \\
0.01   & 2.7e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=2000$ & $\sigma=1.0$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 4.4e-3  \\
10     & 1.3e-3  \\
1      & \textbf{1.1e-3}  \\
0.1    & 2.0e-3  \\
0.01   & 3.6e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=2000$ & $\sigma=1.1$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 6.8e-3  \\
10     & \textbf{4.6e-3}  \\
1      & 7.5e-3  \\
0.1    & 1.3e-2  \\
0.01   & 2.4e-2  \\
\hline
\end{tabular}
%% ---- NEW ROW ----
\vskip 0.2in %% OH this works!
\begin{tabular}{l l}
\hline
$N=4000$ & $\sigma=0.8$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 8.3e-4  \\
10     & 1.3e-4  \\
1      & 2.5e-5  \\
0.1    & \textbf{6.7e-6}  \\
0.01   & 7.4e-6  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=4000$ & $\sigma=0.9$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 1.2e-3  \\
10     & 2.6e-4  \\
1      & \textbf{1.0e-4}  \\
0.1    & 2.0e-4  \\
0.01   & 3.9e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=4000$ & $\sigma=1.0$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 1.9e-3  \\
10     & \textbf{8.9e-4}  \\
1      & 1.6e-3  \\
0.1    & 2.8e-3  \\
0.01   & 5.2e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=4000$ & $\sigma=1.1$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & \textbf{4.3e-3}  \\
10     & 6.0e-3  \\
1      & 1.0e-2  \\
0.1    & 1.2e-2  \\
0.01   & 3.5e-2  \\
\hline
\end{tabular}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

@@

We observe a tradeoff. With smaller $\sigma$, our $C_\sigma$ is closer to the
ideal distribution (as judged by $L_\infty$ error), but this imposes a stricter
upper bound on the sample variance of $\Delta^*$ before our test can be applied,
which thus results in larger minibatch sizes. Conversely, a more liberal upper
bound means we avail ourselves of smaller minibatch sizes, but at the cost of a
less stable derivation for $C_\sigma$.

\textbf{We chose XXXX to use in our experiments, which empirically provided an
acceptable tradeoff to us.} This is reflected in
Algorithm~\ref{alg:our_algorithm}, which assumes that we used \textbf{TODO TODO
TODO TODO TODO} but we reiterate that the choice is arbitrary so long as $0 <
\sigma < \sqrt{\pi^2/3} \approx 1.814$, the standard deviation of the standard
logistic distribution, since there must be some variance left over for $X_{\rm
corr}$.

@@

\begin{table}[t]
\caption{Gaussian Mixture Model Statistics}
\label{tab:poissons}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Metric} & \textbf{Ours} & \textbf{AustereMH} & \textbf{MHSubLhd} \\
\hline \\
Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

>>>>

\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut

@@

% Daniel: use this algorithm style instead
\begin{algorithm}[tb]
\caption{Our acceptance test, {\sc MCminibatch}.}
\label{alg:our_algorithm}
\begin{algorithmic}
    \STATE {\bfseries Input:} number of samples $T$, minibatch size $m$, error
        bound $\delta$, pre-computed correction $C_1(X)$ distribution, initial
        sample $\theta_1$.
    \STATE {\bfseries Output:} a chain of $T$ samples
        $\{\theta_1,\ldots,\theta_T\}$.
    \FOR{$t=1$ {\bfseries to} $T$}
        \STATE -Propose a candidate $\theta'$ from proposal
            $q(\theta' | \theta_t)$.
        \STATE -Draw a minibatch of $m$ points $\{x_1^*,\ldots,x_m^*\}$.
        \STATE -Compute $\Delta^*(\theta_t,\theta')$ and sample variance
            $s^2_{\Delta^*}$.
        \STATE -Estimate moments $\mE[|\Lambda_i-\Lambda|]$ and
            $\mE[|\Lambda_i-\Lambda|^3]$ from the sample, and error $\epsilon$
            from Corollary~\ref{cor:our_bound_delta_prime}.
        \WHILE{$s^2_{\Delta^*} \ge 1$ {\bf or} $\epsilon > \delta$}
            \STATE -Draw $m$ more samples to augment the minibatch, update
                $\Delta^*$, $s^2_{\Delta^*}$ and $\epsilon$ estimates.
        \ENDWHILE
        \STATE -Draw $X_{\rm nc} \sim \mathcal{N}(0,1-s^2_{\Delta^*})$ and
            $X_{\rm corr} \sim C_1(X)$.
    \IF{$\Delta^* + X_{\rm nc} + X_{\rm corr} > 0$}
        \STATE -Accept the candidate, $\theta_{t+1} = \theta'$.
    \ELSE
        \STATE -Reject and re-use the old sample, $\theta_{t+1} = \theta_t$.
    \ENDIF
    \ENDFOR
\end{algorithmic}
\end{algorithm}

With this approach, there is no guarantee that $u^* \geq 0$. However, we have
some flexibility in the choice of $\sigma$ in
Equation~(\ref{eq:overall_corr_problem}).  As we decrease the variance of
$X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount and is
in fact the result of convolution with a Gaussian whose variance is the
difference.  Thus as $\sigma$ decreases, $C_\sigma(X)$ grows and approaches the
derivative of a logistic function at $\sigma = 0$. It retains some weak negative
values for $\sigma > 0$ but removal of those leads to small error. We use
$N=4000$ and $\lambda=10$ for our experiments, which empirically provided
excellent performance. See Table~\ref{tab:xcorr} in
Appendix~\ref{app:correction} for detailed $L_{\rm \infty}$ errors for different
settings. Algorithm~\ref{alg:our_algorithm} describes our procedure, {\sc
MHminibatch}. A few points:

@@

\textbf{Daniel: TODO this should be where we mention that AustereMH still
requires computation of all likelihoods, and we should explain John's complaint
about Gaussianity here.}

@@

In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark our test,
{\sc MHminibatch}, against the test from~\cite{icml2014c1_bardenet14} ({\sc
MHSubLhd}) and from~\cite{cutting_mh_2014}. As highlighted in
Section~\ref{sec:analysis}, the test from~\cite{cutting_mh_2014} can be
implemented using a conservative or a non-conservative variant; we name these
{\sc AustereMH(c)}, and {\sc AustereMH(nc)}, respectively, and they use the
grid-search over minibatch sizes $m$ and per-test thresholds $\epsilon$
described in Appendix~\ref{sssec:grid_search}. Throughout our descriptions, we
refer to a \emph{trial} as the period when an algorithm collects all its desired
samples $\{\theta_1,\ldots,\theta_T\}$, with $T$ usually several thousand.
\textbf{Daniel: better call temperature something other than $T$ ...}

@@

We benchmark with {\sc AustereMH(c)} and {\sc MHSubLhd}. {\sc MHminibatch} and
{\sc MHSubLhd} were initialized with $m=100$. For {\sc MHSubLhd}, we increase
sizes geometrically with $\gamma = 1.5$ and use parameters $p = 2, \delta =
0.01$.  All methods collect 5000 samples using a random walk proposer with
covariance ${\rm diag}(0.15, 0.15)$, which means the M-H test is responsible for
shaping the sample distribution.

Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
samples for the three methods, with darker regions indicating a greater density
of points. There are no obvious differences, showing that {\sc MHminibatch}
reaches an acceptable posterior. We further measure the similarity between each
set of samples and the actual posterior. Due to space constraints, results are
in Appendix~\ref{app:gaussians}.

Figure~\ref{fig:gauss_mix_2} shows that {\sc MHminibatch} dominates in terms of
speed and efficiency. The histograms of the (final) minibatch sizes used each
iteration show that our method consumes significantly less data; the
distribution is short-tailed and the mean is 210, more than an order of
magnitude better compared to the other two methods (averages are 15562 and
16857). The sizes correspond to the running times of the methods, excluding the
likelihood computation of all data points for {\sc MHSubLhd}, which would
drastically increase running time.

@@

% Daniel: moving this here so figures don't overflow to next pages. We can tune
% the positioning if needed.
\begin{figure*}[t]
	\centering
	\includegraphics[width=1.0\linewidth]{logistic_regression_error_regions.png}
	\caption{
    Binary classification accuracy of the MCMC methods on the 1s vs 7s logistic
    regression task for MNIST-13k (left plot) and MNIST-100k (middle plot) as a
    function of cumulative data usage.  The right plot reports performance of
    {\sc MHminibatch} on both datasets when the temperature starts at $T=100$
    and drops to $T=1$ after a ``burn-in'' period of 200 samples of $\theta$.
    For all three plots, one standard deviation is indicated by the shaded error
    regions.
    }
	\label{fig:logistic_performance}
\end{figure*}

\begin{figure*}[t]
	\centering
    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist.png}
	\caption{
    Minibatch sizes for a representative trial of logistic regression on
    MNIST-13k (analogous to Figure~\ref{fig:gauss_mix_2}). Both axes are on a
    log scale and have the same ranges across the three histograms. See
    Section~\ref{ssec:logistic} for details.
    }
	\label{fig:logistic_histograms_mnist}
\end{figure*}

\begin{figure*}[t]
	\centering
    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist8m.png}
	\caption{
    Minibatch sizes for a representative trial of logistic regression on
    MNIST-100k (similar to Figure~\ref{fig:logistic_histograms_mnist}). Both
    axes are on a log scale and have the same ranges across the three
    histograms.  See Section~\ref{ssec:logistic} for details.
    }
	\label{fig:logistic_histograms_mnist8m}
\end{figure*}

\begin{table}[t]
\caption{
Average minibatch sizes ($\pm$ one standard deviation) for the methods on the
logistic regression tasks of binary classification on MNIST-13k and MNIST-100k.
The averages are taken over 10 independent trials (5000 samples each) for
MNIST-13k and 5 independent trials (3000 samples each) for MNIST-100k.
}
\small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
\label{tab:logistic}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Method/Data} & \textbf{MNIST-13k} & \textbf{MNIST-100k}  \\
\hline \\
{\sc MHminibatch}   & $129.8\pm 3.4$    & $202.5 \pm 13.6$ \\
{\sc AustereMH(nc)} & $973.8\pm 49.8$   & N/A \\
{\sc AustereMH(c)}  & $1924.3\pm 52.4$  & $2795.6 \pm 364.0$ \\
{\sc MHSubLhd}      & $10783.4\pm 78.9$ & $14977.3 \pm 582.0$ \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

We next test logistic regression for the binary classification of 1s versus 7s
on the MNIST~\cite{lecun-mnisthandwrittendigit-2010} dataset and (a subset of)
infinite MNIST~\cite{loosli-canu-bottou-2006}. For the former, extracting all 1s
and 7s resulted in 13,000 training samples, and for the latter, we used 87,000
additional (augmented) 1s and 7s to get 100,000 training samples. Both datasets
use the same test set, with 2,163 samples. Henceforth, we call them MNIST-13k
and MNIST-100k, respectively.

For all methods, we impose a uniform prior on $\theta$ and again use a random
walk proposer, with covariance matrix $0.05I$ for MNIST-13k and $0.01I$ for
MNIST-100k. The default temperature setting is a constant at $T=100$ for
MNIST-13k and MNIST-100k. Performance of all methods implicitly relies on the
step size and temperature. Setting temperature too low or step size too high
will generally result in worse performance for all methods. For MNIST-13k, each
method generated 5000 samples for ten independent trials; due to MNIST-100k's
higher computational requirements, the methods generated 3000 samples for five
independent trials.  For additional parameter settings and an investigation on
tuning step sizes, see Appendix~\ref{app:logistic}.

For {\sc MHSubLhd}, we tried to use the provided symbolic bound for
$C_{\theta,\theta'}$ described in~\cite{icml2014c1_bardenet14}, but it was too
high and provided no performance benefit. Instead we use the empirical
$C_{\theta,\theta'}$ from the entire dataset.

The first two subplots of Figure~\ref{fig:logistic_performance} display the
prediction accuracy on both datasets for all methods as a function of the
cumulative training points processed.\footnote{The curves do not span the same
length over the x-axis since the methods consume different amounts of data.} To
generate the curves, for each of the sampled vectors $\theta_t$,
$t\in\{1,\ldots,N\}$ ($N=5000$ for MNIST-13k, $N=3000$ for MNIST-100k), we use
$\theta_t$ as the parameter for logistic regression. The results indicate that
our test is more efficient, obtaining convergence more than an order of
magnitude faster than {\sc AustereMH(nc)} and several orders of magnitude
compared to {\sc AustereMH(c)} and {\sc MHSubLhd}. We also observe the advantage
of having higher temperature from the third plot in
Figure~\ref{fig:logistic_performance}. During the exploration period, the
accuracy rapidly increases, and then after 200 samples, we switch the
temperature to 1 (equivalent to ignoring temperature). {\rm MHminibatch}
requires the step size to decrease, hence the smaller changes in accuracy.

Figures~\ref{fig:logistic_histograms_mnist}
and~\ref{fig:logistic_histograms_mnist8m} show histograms of minibatch sizes for
the three methods on a log-log scale for MNIST-13k and MNIST-100k, respectively.
The histograms only represent one representative trial; Table~\ref{tab:logistic}
contains the average of the average minibatch sizes ($\pm$ one standard
deviation) across either 10 or 5 trials. The results indicate that {\sc
MHminibatch}, with average minibatch sizes of roughly 129.8 and 202.5 for
MNIST-13k and MNIST-100k, respectively, consumes almost an order of magnitude
fewer data points than the next-best method, {\sc AustereMH(nc)}.  We reiterate,
however, that both {\sc AustereMH(nc)} and {\sc MHSubLhd} require computing
$\log p(x_i|\theta)$ and $\log p(x_i|\theta')$ for all $x_i$ each iteration.
These results here do not count that extra data consumption. Only our method and
{\sc AustereMH(c)} rely solely on the minibatch of data each iteration.

@@

\begin{table}[t]
\caption{\textbf{TODO TABLE WITH SMF RESULTS}}
\small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
\label{tab:smf}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{RMSE} & \textbf{MovieLens 10m} & \textbf{Netflix}  \\
\hline \\
{\sc MHminibatch} & ??? & ??? & \\
ADAGrad           & ??? & ??? & \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

@@

We have derived an M-H test for minibatch MCMC which approximates full data
tests. We present theoretical results and experimentally show the benefits of
our test on Gaussian mixtures, logistic regression, and sparse matrix
factorization.  Directions for future work include integrating our algorithm
with~\cite{cutting_mh_2014} by applying both tests each iteration or utilizing
the variance reduction techniques suggested in~\cite{DBLP:conf/icml/ChenG16}.
More elaborate extensions involve running more experiments on other models such
as neural networks and providing a recipe for how to use our algorithm
(following the framework of~\cite{sgmcmc_2015}).

@@

This appendix is divided into two main parts. Appendix~\ref{app:proofs} provides
the proofs that we omitted from the main text due to space constraints.
Appendix~\ref{app:experiments} provides further details on the correction
distribution derivation and on our three main experiments to assist
understanding and reproducibility. Our code is open-source on
GitHub.\footnote{\url{https://github.com/BIDData/BIDMach/blob/master/src/main/scala/BIDMach/updaters/MHTest.scala}}
% Daniel: TODO remove for anonymous review later.

@@

We first observe that
\[
    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
\]
and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
\begin{equation}
-\epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
\end{equation}
as desired.

@@

%% % Daniel: moved to appendix so we can expand it. Originally we had only two
%% % lines, and it looked pretty awkward. So let's add more. Note: this also
%% % depends on whether we use [-10,10] for the X_c range, or [-20,20], etc...
\begin{table}[t]
\caption{Errors ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm
log}$, with $N=4000$ (top row) and $N=2000$ (bottom row).}
\label{tab:xcorr}
\vskip 0.15in
\begin{center}
\begin{small}
%% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\T\B  $N=2000$ & $\sigma=0.8$ \rule{0pt}{3ex} \\
\hline
\T\B $\lambda$ & $L_{\infty}$ error \\
\hline
100    & 2.6e-3 \T \\
10     & 4.0e-4  \\
1      & 6.7e-5  \\
0.1    & 1.4e-5  \\
0.01   & \textbf{5.0e-6} \B \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\T\B $N=2000$ & $\sigma=0.9$ \\
\hline
\T\B $\lambda$ & $L_{\infty}$ error \\
\hline
100    & 3.3e-3 \T \\
10     & 6.4e-4  \\
1      & 1.6e-4  \\
0.1    & \textbf{1.3e-4}  \\
0.01   & 2.7e-4 \B \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\T\B $N=2000$ & $\sigma=1.0$ \\
\hline
\T\B $\lambda$ & $L_{\infty}$ error \\
\hline
100    & 4.4e-3 \T \\
10     & 1.3e-3  \\
1      & \textbf{1.1e-3}  \\
0.1    & 2.0e-3  \\
0.01   & 3.6e-3 \B \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\T\B $N=2000$ & $\sigma=1.1$ \\
\hline
\T\B $\lambda$ & $L_{\infty}$ error \\
\hline
100    & 6.8e-3 \T \\
10     & \textbf{4.6e-3}  \\
1      & 7.5e-3  \\
0.1    & 1.3e-2  \\
0.01   & 2.4e-2 \B \\
\hline
\end{tabular}
%% ---- NEW ROW ----
\vskip 0.2in %% OH this works!
\begin{tabular}{l l}
\hline
\T\B $N=4000$ & $\sigma=0.8$ \\
\hline
\T\B $\lambda$ & $L_{\infty}$ error \\
\hline
100    & 8.3e-4 \T \\
10     & 1.3e-4  \\
1      & 2.5e-5  \\
0.1    & \textbf{6.7e-6}  \\
0.01   & 7.4e-6 \B \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\T\B $N=4000$ & $\sigma=0.9$ \\
\hline
\T\B $\lambda$ & $L_{\infty}$ error \\
\hline
100    & 1.2e-3 \T \\
10     & 2.6e-4  \\
1      & \textbf{1.0e-4}  \\
0.1    & 2.0e-4  \\
0.01   & 3.9e-4 \B \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\T\B $N=4000$ & $\sigma=1.0$ \\
\hline
\T\B $\lambda$ & $L_{\infty}$ error \\
\hline
100    & 1.9e-3 \T \\
10     & \textbf{8.9e-4}  \\
1      & 1.6e-3  \\
0.1    & 2.8e-3  \\
0.01   & 5.2e-3 \B \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\T\B $N=4000$ & $\sigma=1.1$ \\
\hline
\T\B $\lambda$ & $L_{\infty}$ error \\
\hline
100    & \textbf{4.3e-3} \T \\
10     & 6.0e-3  \\
1      & 1.0e-2  \\
0.1    & 1.2e-2  \\
0.01   & 3.5e-2 \B \\
\hline
\end{tabular}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

@@

We observe the expected tradeoff. With smaller $\sigma$, our $C_\sigma$ is
closer to the ideal distribution (as judged by $L_\infty$ error), but this
imposes a stricter upper bound on the sample variance of $\Delta^*$ before our
test can be applied, which thus results in larger minibatch sizes. Conversely, a
more liberal upper bound means we avail ourselves of smaller minibatch sizes,
but at the cost of a less stable derivation for $C_\sigma$.

We chose $N=4000, \sigma=1$, and $\lambda = 10$ to use in our experiments, which
empirically exhibits excellent performance. This is reflected in the description
of {\sc MHminibatch} in Algorithm~\ref{alg:our_algorithm}, which assumes that we
used $\sigma=1$ but we reiterate that the choice is arbitrary so long as $0 <
\sigma < \sqrt{\pi^2/3} \approx 1.814$, the standard deviation of the standard
logistic distribution, since there must be some variance left over for $X_{\rm
corr}$.

@@

\begin{table}[t]
\caption{Gaussian Mixture Model Statistics}
\small
\label{tab:poissons}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Metric} & {\sc MCminibatch} & {\sc AustereMH} & {\sc MHSubLhd} \\
\hline \\
Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

% Daniel: added this here, because it applies to both this and logistic
% regression.
\subsubsection{Grid Search}\label{sssec:grid_search}
